\assignment

### 2017-04-03 project/paper.rnw
Fix cache for presentation.

<<polyfit1, cache = TRUE, dependson = 'sim2000obs', warning = FALSE>>=
# Do a simple fit where spatstat sets up the quadrature so we can see where it
# puts the dummy points.
ex_fit1 <- ppm(ex_ppp ~ x + y)

# Create a quadrature scheme manually to ensure there are dummy points on
# each transect.
ex_Q <- quad(data = ex_ppp, dummy = runifpoint(500000, Window(ex_ppp)))

ex_fit2 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 2), correction = 'iso')
ex_fit3 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 3), correction = 'iso')
ex_fit4 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 4), correction = 'iso')
ex_fit5 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 5), correction = 'iso')
ex_fit6 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 6), correction = 'iso')
ex_fit7 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 7), correction = 'iso')
ex_fit8 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 8), correction = 'iso')
ex_fit9 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 9), correction = 'iso')
ex_fit10 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 10), correction = 'iso')
ex_fit11 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 11), correction = 'iso')
ex_fit12 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 12), correction = 'iso')

# Get predictions and SEs.
ex_p2 <- predict(ex_fit2, window = site_window, se = TRUE)
ex_p3 <- predict(ex_fit3, window = site_window, se = TRUE)
ex_p4 <- predict(ex_fit4, window = site_window, se = TRUE)
ex_p5 <- predict(ex_fit5, window = site_window, se = TRUE)
ex_p6 <- predict(ex_fit6, window = site_window, se = TRUE)
ex_p7 <- predict(ex_fit7, window = site_window, se = TRUE)
ex_p8 <- predict(ex_fit8, window = site_window, se = TRUE)
ex_p9 <- predict(ex_fit9, window = site_window, se = TRUE)
ex_p10 <- predict(ex_fit10, window = site_window, se = TRUE)
ex_p11 <- predict(ex_fit11, window = site_window, se = TRUE)
ex_p12 <- predict(ex_fit12, window = site_window, se = TRUE)

ex_kde <- density(ex_ppp_full)
@

Figure~\ref{fig:polyplot1} shows the predicted intensity surfaces for the
polynomial models, along with a kernel density estimate for comparison. The
2nd-order polynomial surface is too simple to capture the bimodal surface,

The kernel density estimate is computed using the \texttt{spatstat}'s default
``simple rule of thumb'' to select the smoothing bandwidth. Because the
anomalies located more than three feet from a transect were not observed, I
would expect the kernel density estimate to show ridges of high intensity
along the transects. As it turns out, the default bandwith is large enough to
smooth across multiple transects so the ridges do not appear. However, the
kernel density estimate is biased low because not all anomalies are observed;
I divide the estimate by the proportion of the area observed as an approximate
bias correction.

@@

# Multiply predictions by 43560 to convert from anomalies per square foot
# to anomalies per acre.
plot(43560 * ex_p2$estimate, zlim = c(0, 500), main = '2nd Degree')
plot(43560 * ex_p3$estimate, zlim = c(0, 500), main = '3rd Degree')
plot(43560 * ex_p4$estimate, zlim = c(0, 500), main = '4th Degree')
plot(43560 * ex_p5$estimate, zlim = c(0, 500), main = '5th Degree')
plot(43560 * ex_p6$estimate, zlim = c(0, 500), main = '6th Degree')
plot(43560 * ex_p7$estimate, zlim = c(0, 500), main = '7th Degree')
plot(43560 * ex_p8$estimate, zlim = c(0, 500), main = '8th Degree')
plot(43560 * ex_p9$estimate, zlim = c(0, 500), main = '9th Degree')
plot(43560 * ex_p10$estimate, zlim = c(0, 500), main = '10th Degree')
plot(43560 * ex_p11$estimate, zlim = c(0, 500), main = '11th Degree')
plot(43560 * ex_p12$estimate, zlim = c(0, 500), main = '12th Degree')
plot(43560 * ex_kde * area(ex_ppp_full) / area(ex_ppp), zlim = c(0, 500),
     main = 'Kernel Density')
@

<<polyse1, fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = 'Log standard errors and relative standard errors for six of the polynomial models.'>>=
par(mfrow = c(4, 3), mar = c(1, 0, 2, 3), las = 2, cex = 1)
plot(log(43560 * ex_p2$se), main = '2nd Degree\nlog(SE)')
plot(log(43560 * ex_p4$se), main = '4th Degree\nlog(SE)')
plot(log(43560 * ex_p6$se), main = '6th Degree\nlog(SE)')
plot(ex_p2$se / ex_p2$estimate, main = '2nd Degree\nRSE')
plot(ex_p4$se / ex_p4$estimate, main = '4th Degree\nRSE')
plot(ex_p6$se / ex_p6$estimate, main = '6th Degree\nRSE')

plot(log(43560 * ex_p8$se), main = '8th Degree\nlog(SE)')
plot(log(43560 * ex_p10$se), main = '10th Degree\nlog(SE)')
plot(log(43560 * ex_p12$se), main = '12th Degree\nlog(SE)')
plot(ex_p8$se / ex_p8$estimate, main = '8th Degree\nRSE')
plot(ex_p10$se / ex_p10$estimate, main = '10th Degree\nRSE')
plot(ex_p12$se / ex_p12$estimate, main = '12th Degree\nRSE')
@

@@

@@

\appendix
\section{R Code Appendix}

@@

>>>>

<<quad1, cache = TRUE, dependson = 'sim2000obs', warning = FALSE>>=
# Do a simple fit where spatstat sets up the quadrature so we can see where it
# puts the dummy points.
ex_fit1 <- ppm(ex_ppp ~ x + y)

# Create a quadrature scheme manually to ensure there are dummy points on
# each transect.
ex_Q <- quad(data = ex_ppp, dummy = runifpoint(500000, Window(ex_ppp)))
@

<<polyfit2, cache = TRUE, dependson = 'quad1'>>=
ex_fit2 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 2), correction = 'iso')
@
<<polyfit3, cache = TRUE, dependson = 'quad1'>>=
ex_fit3 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 3), correction = 'iso')
@
<<polyfit4, cache = TRUE, dependson = 'quad1'>>=
ex_fit4 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 4), correction = 'iso')
@
<<polyfit5, cache = TRUE, dependson = 'quad1'>>=
ex_fit5 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 5), correction = 'iso')
@
<<polyfit6, cache = TRUE, dependson = 'quad1'>>=
ex_fit6 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 6), correction = 'iso')
@
<<polyfit7, cache = TRUE, dependson = 'quad1'>>=
ex_fit7 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 7), correction = 'iso')
@
<<polyfit8, cache = TRUE, dependson = 'quad1'>>=
ex_fit8 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 8), correction = 'iso')
@
<<polyfit9, cache = TRUE, dependson = 'quad1'>>=
ex_fit9 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 9), correction = 'iso')
@
<<polyfit10, cache = TRUE, dependson = 'quad1'>>=
ex_fit10 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 10), correction = 'iso')
@
<<polyfit11, cache = TRUE, dependson = 'quad1'>>=
ex_fit11 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 11), correction = 'iso')
@
<<polyfit12, cache = TRUE, dependson = 'quad1'>>=
ex_fit12 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 12), correction = 'iso')
@

<<polypred1>>=
# Get predictions and SEs.
ex_p2 <- predict(ex_fit2, window = site_window, se = TRUE)
ex_p3 <- predict(ex_fit3, window = site_window, se = TRUE)
ex_p4 <- predict(ex_fit4, window = site_window, se = TRUE)
ex_p5 <- predict(ex_fit5, window = site_window, se = TRUE)
ex_p6 <- predict(ex_fit6, window = site_window, se = TRUE)
ex_p7 <- predict(ex_fit7, window = site_window, se = TRUE)
ex_p8 <- predict(ex_fit8, window = site_window, se = TRUE)
ex_p9 <- predict(ex_fit9, window = site_window, se = TRUE)
ex_p10 <- predict(ex_fit10, window = site_window, se = TRUE)
ex_p11 <- predict(ex_fit11, window = site_window, se = TRUE)
ex_p12 <- predict(ex_fit12, window = site_window, se = TRUE)

ex_kde <- density(ex_ppp_full) * area(ex_ppp_full) / area(ex_ppp)
@

Figure~\ref{fig:polyplot1} shows the predicted intensity surfaces for the
polynomial models, along with a kernel density estimate for comparison.
500,000 dummy points do not seem to be enough for the likelihood
approximation to converge to the Poisson likelihood, but the predicted
intensity values are reasonable (discussed further in the next section).
Because of the lack of convergence, I do not do any formal model comparisons.

The second- and third-order polynomial models are too simple to capture the
bimodality of the true surface, but the fourth- and higher-order models have
large, high-intensity regions that coincide with both concentrated munitions
use regions of the true surface. Up until degree nine, the polynomial models
are smoother than the kernel intensity estimate. The ninth-, tenth-, and
eleventh-degree models all look similar to the kernel intensity estimate,
while the twelfth-degree model is noticeably noisier. Edge effects are
noticeable for \(p \geq 4\) with some predictions along the boundary abruptly
dropping to zero or increasing to hundreds of anomalies per acre higher than
any of the predictions in the interior of the region.

The kernel density estimate is computed using the \texttt{spatstat}'s default
``simple rule of thumb'' to select the smoothing bandwidth. Because the
anomalies located more than three feet from a transect were not observed, I
would expect the kernel density estimate to show ridges of high intensity
along the transects. As it turns out, the default bandwith is large enough to
smooth across multiple transects so the ridges do not appear. However, the
kernel density estimate is biased low because not all anomalies are observed;
I divide the estimate by the proportion of the area observed as an approximate
bias correction and the result is a surprisingly reasonable intensity map.

@@

# Multiply predictions by 43560 to convert from anomalies per square foot
# to anomalies per acre.
plot(43560 * ex_p2$estimate, zlim = c(0, 300), main = '2nd Degree')
plot(43560 * ex_p3$estimate, zlim = c(0, 300), main = '3rd Degree')
plot(43560 * ex_p4$estimate, zlim = c(0, 300), main = '4th Degree')
plot(43560 * ex_p5$estimate, zlim = c(0, 300), main = '5th Degree')
plot(43560 * ex_p6$estimate, zlim = c(0, 300), main = '6th Degree')
plot(43560 * ex_p7$estimate, zlim = c(0, 300), main = '7th Degree')
plot(43560 * ex_p8$estimate, zlim = c(0, 300), main = '8th Degree')
plot(43560 * ex_p9$estimate, zlim = c(0, 300), main = '9th Degree')
plot(43560 * ex_p10$estimate, zlim = c(0, 300), main = '10th Degree')
plot(43560 * ex_p11$estimate, zlim = c(0, 300), main = '11th Degree')
plot(43560 * ex_p12$estimate, zlim = c(0, 300), main = '12th Degree')
plot(43560 * ex_kde, zlim = c(0, 500), main = 'Kernel Density')
@

Figure~\ref{fig:polyse1} shows the log-scale standard errors for six of the
polynomial models. For all models except the twelfth-degree model, the
standard error is low across most of the site but high along the boundary. The
standard errors increase with the degree of the model. This is expected and
resonable because locations in the interior of the region have the most nearby
anomalies and dummy points, and predictions from more complex models depend on
more estimated parameters.

If the true number of high-intensity regions was known to be 2, any polynomial
model with \(p \geq 4\) (or the kernel density estimate) would provide an
accurate (if imprecise) indication of where to focus the UXO remediation
effort. However, the plots alone provide little indication of which model has
the ideal level of complexity, but there are several practical difficulties
with fitting these models that should be sorted out before I would be
comfortable making recommendations about how to select an optimal model for
prediction outside the observed subregion.

<<polyse1, fig.width = 6, fig.height = 4, fig.pos = 't', fig.cap = 'Log standard errors for six of the polynomial models.'>>=
par(mfrow = c(2, 3), mar = c(1, 0, 2, 3), las = 2, cex = 1)
plot(log(43560 * ex_p2$se), main = '2nd Degree\nlog(SE)')
plot(log(43560 * ex_p4$se), main = '4th Degree\nlog(SE)')
plot(log(43560 * ex_p6$se), main = '6th Degree\nlog(SE)')
plot(log(43560 * ex_p8$se), main = '8th Degree\nlog(SE)')
plot(log(43560 * ex_p10$se), main = '10th Degree\nlog(SE)')
plot(log(43560 * ex_p12$se), main = '12th Degree\nlog(SE)')
@

@@

<<dummy>>=
par(mfrow = c(1, 2), mar = c(1, 0, 1, 0), cex = 1)
plot(ex_fit1$Q$dummy, pch = '.', border = NA,
     main = 'Default Dummy Points')
plot(site_window, add = TRUE)
plot(ex_Q$dummy, pch = '.', border = NA,
     main = 'Manual Dummy Points')
plot(site_window, add = TRUE)
@

@@

@@

\pagebreak
\appendix
\section*{R Code Appendix}

<<packages, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<hw4_sim, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<hw4_fit_full, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<hw4_fit_s, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<hw4_plot, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<hw4_fitplot, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<uxoexampledata, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<truth, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<sim2000obs, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<quad1, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<polyfit2, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit3, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit4, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit5, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit6, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit7, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit8, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit9, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit10, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit11, echo = TRUE, eval = FALSE>>=
@
\vspace{-30pt}
<<polyfit12, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<polypred1, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<polyplot1, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<polyse1, echo = TRUE, eval = FALSE>>=
@
\vspace{-20pt}
<<dummy, echo = TRUE, eval = FALSE>>=
@
